---
title: "Homework5"
output: github_document
---

First load the packages

```{r}
library(gapminder)
library(tidyverse)
library(forcats)
library(plotly)
```

# Factor Management and Faceting

First lets explore how to partially (and later completely) remove Oceania from the *Gapminder* dataset. The `summary` function breaks down the `continent` variable to show the frequency at which each data point occurs per continent.

```{r}
summary(gapminder$continent)
```

There are five continents, and 24 entries belong to Oceania. Removing one continent leaves us with four, so `four_cont` can be our variable name. We can use `summary` on the `four_cont` dataset to check if the countries that belong to Oceania have been removed.

```{r}
four_cont <- gapminder %>%
  filter(continent != "Oceania")
summary(four_cont$continent)
```

However, completely removing trace that Oceania existed requires the `droplevels` function. To keep thing organized, this would require a new variable name that builds upon the last one used, like `no_oceania`.

```{r}
no_oceania <- droplevels(four_cont)
summary(no_oceania$continent)
```

Now lets look at the median and lowest lifeExp of countries within the Americas continent, sorted ascendingly by lowest lifeExp.

```{r}
americaslifeExp <- gapminder %>%
  filter(continent == "Americas") %>%
  group_by(country) %>%
  summarize(
    Continent = unique(continent),
    MinLifeExp = min(lifeExp),
    MedianLifeExp = median(lifeExp)
  ) %>%
  arrange(MinLifeExp)
knitr::kable(americaslifeExp)
```

We can reorder for the dataset to show descending MinLifeExp.

```{r}
americaslifeExp %>%
  arrange(desc(MinLifeExp)) %>%
  knitr::kable()
```

Now lets take the `no_oceania` dataset and plot it for 1992 with faceting for lifeExp.

```{r}
no_oceania %>%
  filter(year == 1992) %>%
  mutate(continent = fct_reorder(continent, lifeExp)) %>%
  ggplot(aes(continent, lifeExp)) +
  geom_violin(aes()) +
  theme(
    plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1),
    axis.line = element_line(
      size = 0.3,
      linetype = "solid"
    ),
    axis.ticks = element_line(size = 0.3),
    axis.title = element_text(family = "mono"),
    axis.text = element_text(family = "mono"),
    axis.text.x = element_text(family = "mono"),
    axis.text.y = element_text(family = "mono"),
    panel.background = element_rect(fill = NA)
  ) +
  labs(x = "Continent", y = "Life Expectancy in 1992")
```

The faceting can be modified to tell different stories, here continents are sorted descendingly by median, which is where the "broadest" region of the violin plot exists. This can 

```{r}
no_oceania %>%
  filter(year == 1992) %>%
  mutate(continent = fct_reorder(continent, lifeExp, median, .desc = TRUE)) %>%
  ggplot(aes(continent, lifeExp)) +
  geom_violin(aes()) +
  theme(
    plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1),
    axis.line = element_line(
      size = 0.3,
      linetype = "solid"
    ),
    axis.ticks = element_line(size = 0.3),
    axis.title = element_text(family = "mono"),
    axis.text = element_text(family = "mono"),
    axis.text.x = element_text(family = "mono"),
    axis.text.y = element_text(family = "mono"),
    panel.background = element_rect(fill = NA)
  ) +
  labs(x = "Continent", y = "Life Expectancy in 1992")
```

The same can be done for the lifeExp of each country (not including Oceania)

```{r}
no_oceania %>%
  filter(year == 1992) %>%
  mutate(country = fct_reorder(country, lifeExp)) %>%
  ggplot(aes(country, lifeExp)) +
  geom_point(aes()) +
  theme(
    plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1),
    axis.line = element_line(
      size = 0.3,
      linetype = "solid"
    ),
    axis.ticks = element_line(size = 0.3),
    axis.title = element_text(family = "mono"),
    axis.text = element_text(family = "mono"),
    axis.text.x = element_text(family = "mono"),
    axis.text.y = element_text(family = "mono"),
    panel.background = element_rect(fill = NA)
  ) +
  labs(x = "Country", y = "Life Expectancy in 1992") +
  theme(axis.text = element_text(
    hjust = 1,
    vjust = 0
  ), axis.text.x = element_text(
    colour = "black",
    vjust = 0.4, angle = 90
  )) + theme(axis.text.x = element_text(size = 5)) +
  theme(plot.background = element_rect(size = 1))
```

Then faceted to be sorted descendingly according to lifeExp.

```{r}
no_oceania %>%
  filter(year == 1992) %>%
  mutate(country = fct_reorder(country, lifeExp, .desc = TRUE)) %>%
  ggplot(aes(country, lifeExp)) +
  geom_point(aes()) +
  theme(
    plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1),
    axis.line = element_line(
      size = 0.3,
      linetype = "solid"
    ),
    axis.ticks = element_line(size = 0.3),
    axis.title = element_text(family = "mono"),
    axis.text = element_text(family = "mono"),
    axis.text.x = element_text(family = "mono"),
    axis.text.y = element_text(family = "mono"),
    panel.background = element_rect(fill = NA)
  ) +
  labs(x = "Country", y = "Life Expectancy in 1992") +
  theme(axis.text = element_text(
    hjust = 1,
    vjust = 0
  ), axis.text.x = element_text(
    colour = "black",
    vjust = 0.4, angle = 90
  )) + theme(axis.text.x = element_text(size = 5)) +
  theme(plot.background = element_rect(size = 1))
```

# Reading and Writing CSV files

The `no_oceania` dataset can be exported, saving the output of transformations done with R.

```{r}
write_csv(no_oceania, "GapminderWithoutOceania.csv")
```

This file can be read from the working directory

```{r}
read_csv("GapminderWithoutOceania.csv")
```

# Remaking a plot

Here a plot shown before is slightly modified to show a color dimension for continent. The plot elements, like the color legend, is edited for optimal use of space.

```{r}
no_oceania %>%
  filter(year == 1992) %>%
  mutate(country = fct_reorder(country, lifeExp)) %>%
  ggplot(aes(country, lifeExp)) +
  geom_point(aes(color = continent)) +
  theme(
    plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1),
    axis.line = element_line(
      size = 0.3,
      linetype = "solid"
    ),
    axis.ticks = element_line(size = 0.3),
    axis.title = element_text(family = "mono"),
    axis.text = element_text(family = "mono"),
    axis.text.x = element_text(family = "mono"),
    axis.text.y = element_text(family = "mono"),
    panel.background = element_rect(fill = NA)
  ) +
  labs(x = "Country", y = "Life Expectancy in 1992") +
  theme(axis.text = element_text(
    hjust = 1,
    vjust = 0
  ), axis.text.x = element_text(
    colour = "black",
    vjust = 0.4, angle = 90
  )) + theme(axis.text.x = element_text(size = 5)) +
  theme(plot.background = element_rect(size = 1)) + theme(
    legend.text = element_text(family = "mono"),
    legend.title = element_text(family = "mono"),
    legend.position = "top", legend.direction = "horizontal"
  ) + labs(colour = NULL)
```

To make this graph interactive in an HTML setting, the `plotly` library can enable interactive properties. However, much of the custom edits available in the `ggplot2` package are ignored by `plotly`. The HTML output is commented out (suppressed) because it interferes with .md kitting into github_document format.

```{r}
plotly_plot <- no_oceania %>%
  filter(year == 1992) %>%
  mutate(country = fct_reorder(country, lifeExp)) %>%
  ggplot(aes(country, lifeExp)) +
  geom_point(aes(color = continent)) +
  theme(
    plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1),
    axis.line = element_line(
      size = 0.3,
      linetype = "solid"
    ),
    axis.ticks = element_line(size = 0.3),
    axis.title = element_text(family = "mono"),
    axis.text = element_text(family = "mono"),
    axis.text.x = element_text(family = "mono"),
    axis.text.y = element_text(family = "mono"),
    panel.background = element_rect(fill = NA)
  ) +
  labs(x = "Country", y = "Life Expectancy in 1992") +
  theme(axis.text = element_text(
    hjust = 1,
    vjust = 0
  ), axis.text.x = element_text(
    colour = "black",
    vjust = 0.4, angle = 90
  )) + theme(axis.text.x = element_text(size = 5)) +
  theme(plot.background = element_rect(size = 1)) + theme(
    legend.text = element_text(family = "mono"),
    legend.title = element_text(family = "mono"),
    legend.position = "top", legend.direction = "horizontal"
  ) + labs(colour = NULL)

# ggplotly(plotly_plot)
```

# Saving images

Here the plot shown above is exported as a high-resolution .png image.

```{r}
ggsave("NoOceania.png",
  plot = plotly_plot, 
  device = "png",
  width = 12, 
  height = 6,
  dpi = 500)
```

Then it can be called back. 
